---
title: "MitoFinder figures"
author: "Mareike"
date: "10/13/2020"
output: html_document
---

Scripts for plotting how well MitoFinder worked

Load packages
```{r}
library(ggplot2)
library(tidyr)
library(ggpubr)
library(viridis)
library(cowplot)
```

Import results file
```{r}
mito_df <- read.delim("../Analyses/MitoFinder/output/mitoFinder_results_final_length_cov_MolEcol.txt", header = TRUE, sep = "\t")

#remove sequences from samples that were sequenced twice
mito_df <- subset(mito_df, mito_df$Sample != "PD_0324" & mito_df$Sample != "PD_0382" & mito_df$Sample != "PD_0369" & mito_df$Sample != "PD_0297")

#mito_df <- mito_df %>% drop_na()
mito_df$Assembler <- "metaSPAdes"
mito_df$Annotator <- "MiTFi"
```

Import results file for samples with multiple contigs
```{r}
multi_df <- read.delim("../Analyses/MitoFinder/output/mitoFinder_results_multiple_contigs_all.txt", header = TRUE, sep = "\t")

#remove sequences from samples that were sequenced twice
multi_df <- subset(multi_df, multi_df$Sample != "PD_0324" & multi_df$Sample != "PD_0382" & multi_df$Sample != "PD_0369" & multi_df$Sample != "PD_0297")
```

Stats
```{r}
summary(subset(mito_df, mito_df$Sample != "PD_0084"))
summary(mito_df)
summary(subset(multi_df, multi_df$ContigNum != "contig_1"))
summary(subset(multi_df, multi_df$ContigNum == "contig_1"))
```


Plot number of contigs found:

```{r}
p_contigs <- ggplot(mito_df, aes(y =ContigsFound, x= Assembler, fill = Assembler)) + 
  theme_classic() +
#  geom_violin(alpha=0.7) +
  geom_jitter(aes(Assembler), height = 0, width = 0.2, pch = 16, alpha = 0.6) +
  geom_violin(alpha = 0.4, trim = FALSE) +
#  geom_beeswarm(alpha = 0.5, side = -1L) +
  xlab(NULL) +
  ylab("Number of contigs found") +
  scale_fill_manual(values = "#0D0887FF") +
  scale_color_manual(values = "#0D0887FF") +
  guides(fill = "none", color = "none") +
  scale_y_continuous(breaks = seq(0, 6, len = 7)) +
#  ggtitle("Contigs") +
  theme(plot.title = element_text(face = "bold"),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

p_contigs
```

Plot length of final contig:
```{r}
p_length <- ggplot(mito_df, aes(x = Assembler, y = LengthFinalContig/1000, fill = Assembler)) +
  geom_violin(alpha = 0.5,trim = FALSE) +
  geom_jitter(height = 0, pch = 16, alpha = 0.6) + 
  scale_fill_manual(values = "#6A00A8FF") +
  guides(fill = "none") +
  xlab(NULL) +
  ylab("Length of final contig (kb)") +
  theme_classic() +
#  ggtitle("Length") +
  ylim(15.3, 16.9) +
  theme(plot.title = element_text(face = "bold"),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

p_length
```

Plot number of genes annotated:
```{r}
p_genes <- ggplot(mito_df, aes(x = Annotator, y = GenesAnnotated, fill = Annotator)) +
#  geom_violin(alpha = 0.5) +
  geom_jitter(height = 0, pch = 16, alpha = 0.6) +
  guides(fill = F) +
  xlab(NULL) +
  ylab("Number of genes annotated") +
  ylim(0,17) +
  theme_classic() +
#  ggtitle("Genes") +
  theme(plot.title = element_text(face = "bold"),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

p_genes
```

Plot length of contigs when multiple contigs found:
```{r}
p_contigsLength <- ggplot(multi_df, aes(x = ContigNum, y = ContigLength/1000, fill = ContigNum)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_jitter(width = 0.1, pch = 16, alpha = 0.6) +
  theme_classic() +
  scale_fill_manual(values = viridis(6, option = "C")) +
  guides(fill = "none") +
  xlab("Contig") +
  ylab("Length (kb)") +
#  ggtitle("Length") +
  scale_x_discrete(labels = c("1","2", "3", "4", "5", "6")) +
  theme(plot.title = element_text(face = "bold"))

p_contigsLength
```
Plot coverage of contigs when multiple contigs found:
```{r}
p_contigsCov <- ggplot(multi_df, aes(x = ContigNum, y = ContigCov, fill = ContigNum)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_jitter(width = 0.1, pch = 16, alpha = 0.6) +
  theme_classic() +
  scale_fill_manual(values = viridis(6, option = "C")) +
  guides(fill = "none") +
  xlab("Contig") +
  ylab("Coverage (x)") +
  ylim(0, 1600) +
#  ggtitle("Coverage") +
  scale_x_discrete(labels = c("1","2", "3", "4", "5", "6")) +
  theme(plot.title = element_text(face = "bold"))

p_contigsCov
```

Arrange panels and save:
```{r}
p_combo <- plot_grid(p_contigs, p_length, p_contigsLength, p_contigsCov, nrow = 2, ncol = 2, labels = c("(A)", "(B)", "(C)", "(D)"), align = "v")

#save
ggsave(plot = p_combo, file = "./results_MitoFinder.pdf", height = 6, width = 8, useDingbats= FALSE)

```
